---
title: "Exploring cell population heterogeneity"
author: "Denis Puthier & Jessica Chevallier"
date: "2024-02-07"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteEngine{knitr::knitr}
  %\VignetteIndexEntry{Exploring cell population heterogeneity}
  %\usepackage[UTF-8]{inputenc}
  html_document:
  fig_caption: yes
  highlight: zenburn
  theme: cerulean
  toc: no
  toc_depth: 3
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>", 
  fig.align = "center",
  dpi = 150, 
  cache.rebuild = TRUE
)
```

<style>
body {
text-align: justify}
img {
border: 0;}
</style>

# Objective

The STarlight package is an R package designed for the exploratory analysis and reporting of cell or molecule coordinates obtained from Spatial omics experiments. Here we will explore a subset of the CODEX data from Hickey *et al* (). In their article entitled 'Strategies for Accurate Cell Type Identification in CODEX Multiplexed Imaging Data', Hickey *et al* generated single-cell multiplexed imaging datasets by conducting CODEX on four human colon sections (ascending, transverse, descending, and sigmoid) from the same donor using a panel of 47 oligonucleotide-barcoded antibodies. The regions are therafter refered as regions 1 to 4 respectively. Following cell segmentation, they applied normalization combined with unsupervised clustering algorithms to produce identify cell type populations.
In this vignette we will examplify the use of STarlight to create a Bookdown report based on x/y coordinates of cell populations. 

From the authors: "
Multiplexed imaging techniques allow imaging up to 60 markers in a tissue simultaneously, which increases the number of identifiable cell types in situ [...] ... we created single-cell multiplexed imaging datasets by performing CODEX on four sections of the human colon (ascending, transverse, descending, and sigmoid) using a panel of 47 oligonucleotide-barcoded antibodies. "

((https://pmc.ncbi.nlm.nih.gov/articles/PMC8415085/
https://datadryad.org/dataset/doi:10.5061/dryad.dfn2z352c))

OR THIS ONE

https://datadryad.org/dataset/doi:10.5061/dryad.pk0p2ngrf 
B004_Ascending

scan("~/Downloads/cell_1_annot.csv", what = "character", n = 10)

```{r}
library(STarlight)

st_file <- file.path("/Users/puthier/Documents/ST_data/codex_Hickey/23_09_CODEX_HuBMAP_alldata_Dryad_merged.csv")
st_dt <- as.data.frame(data.table::fread(st_file,
                                          sep = ",",
                                          head = TRUE,
                                          nThread = 4))
uniq_region <- unique(st_dt$unique_region)

mapping <- c("x"="x", "y"="y", feature="Cell Type")

list_st <- list()
for(region in uniq_region){
  print(paste0("Reading ", region))
    list_st[[region]] <- load_spatial(st_dt, 
                                      mapping = mapping, 
                                      sep=",", 
                                      bin_size = 50, 
                                      constrain = list(paste0("unique_region == '", region, "'")))
}

```


```{r}
st_report(list_st, 
          hc_tree_params = list(class_nb = 10, offset = 8))
```



