---
title: "Guided tutorial"
author: "D. Puthier"
date: "2024-03-06"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteEngine{knitr::knitr}
  %\VignetteIndexEntry{Guided tutorial}
  %\usepackage[UTF-8]{inputenc}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```


```{r loadlib}
library(magrittr)
library(Seurat)
library(STarlight)
STarlight::set_verb_level(0)
library(patchwork)
library(ggplot2)
library(dplyr)


```

# About

Spatial transcriptomics enables the spatially resolved measurement of gene expression within tissue sections, providing valuable insights into the spatial organization of biological processes. Assigning RNA molecules to cells using a segmentation process is generally the first step in the analysis. This approach can be tricky for a number of reasons, including high cell density, the absence or poor quality of surface labelling, etc. The `STarlight` package is intended to perform analysis of RNA molecules using a grid-based approach (sometimes referred as 'rasterization'). The `load_spatial()`is the entry point of the `STarlight` package It will load the molecule coordinates and create a 2D binned grid with default size 25µm.


# Retrieving the dataset

Here, for demonstration purpose, we are using a subset (64 genes) of a 10x Genomics dataset tiny dataset used for Xenium Explorer Demo. The analyzed region corresponds to a fresh frozen section of the mouse Brain. The original dataset can be downloaded from [here](https://www.10xgenomics.com/resources/datasets/fresh-frozen-mouse-brain-for-xenium-explorer-demo-1-standard). 


```{r}
## create a temporary folder
tmpdir <- tempdir()
dir.create(tmpdir, showWarnings = FALSE)
remote_file <- "Xenium_Mouse_Brain_Coronal_tiny.csv.gz"
## Download and unzip the file
url_data <- paste0("https://zenodo.org/records/11371819/files/",
                   remote_file)
dest_file <- file.path(tmpdir, remote_file)
download.file(url_data, destfile = dest_file)

```

# Loading data

Here we load the data considering that the Xenium dataset is a simple csv flat file. If the full result directory is available you may use method="xenium" and provide the path to the directory as argument to 'path'.

```{r loadfile}

st_obj <- load_spatial(path = dest_file, 
             method = "coordinates", 
             sep=",", 
             mapping = c(feature="feature_name", 
                         x="x_location", 
                         y="y_location"), 
                         bin_size = 25, 
            control = "(NegControl)|(^BLANK)", verbose = FALSE)

##
```


# The STGrid object

The Spatial Transcriptomic Grid class (STGrid), serves as a fundamental structure for representing spatial transcriptomic analysis results. It provides a structured framework for managing and analyzing spatial transcriptomic data, facilitating comprehensive exploration and interpretation of spatial gene expression patterns within tissue sections.

```{r show, echo = TRUE, eval=TRUE, warning=FALSE, results='hide'}
STarlight::set_verb_level(1)
st_obj
STarlight::set_verb_level(0)
```
```{r summary}
summary(st_obj)
```

```{r slots, echo = TRUE, eval=TRUE, warning=FALSE, results='hide'}
slotNames(st_obj)
```

```{r show_st_methods, echo = TRUE, eval=TRUE, warning=FALSE, results='hide'}
STarlight::show_st_methods()
```

## Subsetting

The subsetting function, implemented as a method for the "[" operator in the STGrid class, provides a versatile means of extracting subsets of data from an STGrid object. This function is designed to facilitate the extraction of specific genes or regions within the spatial transcriptomic data, allowing users to focus their analysis on relevant subsets of the dataset.

### Subsetting using bin_x / bin_y

We can explore the binning information stored within our `st_obj` STGrid object using the `bin_x()` and `bin_y()` methods. Let's take a look at the first few entries of the bin_x and bin_y vectors:

```{r binx_1, echo = TRUE, eval=TRUE, warning=FALSE, results='hide'}
head(bin_x(st_obj))
```

This command retrieves the first few entries of the bin_x vector, providing insights into the bins along the x-axis.

```{r binx_2, echo = TRUE, eval=TRUE, warning=FALSE, results='hide'}
head(bin_y(st_obj))
```

Similarly, this command retrieves the first few entries of the `bin_y` vector, offering insights into the bins along the y-axis. Now, let's leverage the subsetting function to extract data corresponding to specific bins along the x-axis. Here, we retrieve data associated with bins ranging from the 50th to the 100th entry of the bin_x vector:

```{r binx_3, echo = TRUE, eval=TRUE, warning=FALSE, results='hide'}
st_obj[bin_x(st_obj)[50:100], ]
```

This command retrieves data from the STGrid object `st_obj` corresponding to the specified range of bins along the x-axis, providing a subset of spatial transcriptomic data for further analysis.

### Subsetting using genes

We can retrieve spatial transcriptomic data associated with specific genes from our st_obj STGrid object using the subsetting operator. Here we extract data corresponding to the "Rag1" and "Cd3e" genes:

```{r genes, echo = TRUE, eval=TRUE, warning=FALSE, results='hide'}
st_obj[c("Chat", "Cbln4"), ]
```

We may also remove all control probes from the STGrid object to focus our analysis on genes with meaningful expression levels.

```{r delete, echo = TRUE, eval=TRUE, warning=FALSE, results='hide'}
st_obj <- rm_controls(st_obj)
```

### Subsetting using numeric

We can retrieve spatial transcriptomic data based on the position of genes within the `feat_names()` vector of our `st_obj` STGrid object. Let's extract data for the first 10 genes according to their position in the feat_names() vector:

```{r numerics_1, echo = TRUE, eval=TRUE, warning=FALSE, results='hide'}
st_obj[feat_names(st_obj)[1:10], ]
```

Next, let's explore data for the 10th to 12th genes using numeric indices. To begin, let's determine the total number of genes in our STGrid object:

```{r numerics_2, echo = TRUE, eval=TRUE, warning=FALSE, results='hide'}
nb_feat(st_obj)
```

With this information, we can retrieve data for the 10th to 12th genes:

```{r index_10_12, echo = TRUE, eval=TRUE, warning=FALSE, results='hide'}
st_obj[10:12, ]
```

This command retrieves spatial transcriptomic data for genes positioned 10 through 12 in the `feat_names()` vector, facilitating targeted analysis of specific gene subsets.

## Image Visualization

### The spatial_image() function

The `spatial_image()` function allows for the visualization of the density of spatial molecules across the tissue section. The following examples demonstrate the visualization of a single gene. The grid=TRUE argument allow to show bin numbering in x and y dimension. This eases the selection of particular regions.

```{r single_gene, warning=FALSE, results='hide', echo = FALSE, eval=TRUE}
spatial_image(st_obj, feat="Ano1", grid = 20, 
              logb = 10, scale = TRUE, saturation = 0.9)
```

One can **zoom** for instance in the region with high "Ano1" molecule density in the upper right corner of the previous image. Here we use `rebin()` to increase the resolution specifically on "Ano1".

```{r zoom}
x_bins <-  bin_x(st_obj)[181:nbin_x(st_obj)]
y_bins <-  bin_y(st_obj)[101:nbin_y(st_obj)]
st_obj_r1 <- st_obj[x_bins, y_bins]
spatial_image(re_bin(st_obj_r1[c("Ano1", "Chat"),], bin_size = 5, verbose = FALSE), 
              feat=c("Ano1", "Chat"), logb = 10,
              scale = TRUE, saturation = 0.9)
```

### The spatial_plot() function

The `spatial_plot()`function is intented to create a scatter plot of molecule x/y coordinates. We may for example highlight other molecules with same pattern as "Ano1"

```{r spatial_plot}
spatial_plot(st_obj[x_bins, y_bins], feat_list = c("Ano1", 
                                                     "Chat",
                                                     "Ebf3"), 
             colors = ggsci::pal_aaas()(3),
             size=1)
```

# Searching modules

The hc_tree() allow to organize genes into subclusters.

```{r spatial_plot_2}
hc_st <- hc_tree(st_obj, method = "ward.D", 
                 layout = "circular", dist_method = "pearson",
                 class_nb = 20)
hc_st
```

```{r}
hc_st$tree_classes
```

Computing average module score:

```{r}
st_obj <- compute_module_score(st_obj, modules = hc_st$tree_classes)
meta_names(st_obj)
```
```{r}
spatial_image(st_obj, features = meta_names(st_obj))
```

```{r}
spatial_image(st_obj, features = hc_st$tree_classes[["MOD07"]])
```


# Molecule clustering

## K Ripley's function

We may use the K-Ripley's function to assess the level of clustering of each molecules in the two regions. Not that this step is rather long as the density of molecule increases. For the tutorial will thus concentrate on few genes.

```{r compute_k_ripley}
gene_list <- c("Chat", "Spag16", "Tacr1", "Sncg", feat_names(st_obj)[1:10])
plot_rip_k(compute_k_ripley(st_obj, rmax = 80, verbose = FALSE))
```


```{r}
spatial_image(st_obj, features = c("Chat", "Spag16", "Tacr1", "Nwd2"))
```

# Comparing counts

The stcompr() function is used to compare two sets of data. By way of illustration, and to limit the calculation time required for this vignette, we will use this function to compare two regions of the organ, which will be considered here as two conditions. Still for illustration purpose, we will generate two replicates per conditions/regions (in fact two close regions)


```{r STarlight}
## "Conditions 1"
# "replicate 1"
x_bins <-  bin_x(st_obj)[180:nbin_x(st_obj)]
y_bins <-  bin_y(st_obj)[100:nbin_y(st_obj)]
st_obj_r1_1 <- st_obj[x_bins, y_bins]

# "replicate 2"
x_bins <-  bin_x(st_obj)[170:200]
y_bins <-  bin_y(st_obj)[90:130]
st_obj_r1_2 <- st_obj[x_bins, y_bins]


## "Conditions 2"
# "replicate 1"
x_bins <-  bin_x(st_obj)[60:100]
y_bins <-  bin_y(st_obj)[100:nbin_y(st_obj)]
st_obj_r2_1 <- st_obj[x_bins, y_bins]

# "replicate 2"
x_bins <-  bin_x(st_obj)[50:90]
y_bins <-  bin_y(st_obj)[80:130]
st_obj_r2_2 <- st_obj[x_bins, y_bins]
```

Here is our illustrative dataset visualized using the *cmp_images()* function:

```{r}
cmp_images(st_obj_r1_1,
           st_obj_r1_2,
           st_obj_r2_1, 
           st_obj_r2_2,
           feat_list="count_sum",
           names=c("region_1_1", "region_1_2", "region_2_1", "region_2_2"),
           color_y_strip=c("pink", "pink", "lightblue", "lightblue"),
           color_strip_text_y = "black",
           color_x_strip="#333333",
           colors = viridis::magma(10)
           )
```
And the count distribution for each replicate:

```{r}
dist_st(st_obj_r1_1,
           st_obj_r1_2,
           st_obj_r2_1, 
           st_obj_r2_2,
           transform = "log10")
```


## Creating an STCompR object

Now we can create an STCompR object using the stcompr() function. 

```{r STarlight_create}
cmp <- stcompr(list(st_obj_r1_1,
                    st_obj_r1_2),
                list(st_obj_r2_1, 
                    st_obj_r2_2), 
               name_1="region_1", name_2="region_2")
```

## Count distributions

Here are the normalized counts:

```{r cmp_boxplot}
cmp_boxplot(cmp, transform = "log10", normalized = TRUE)
```

## Count comparison

As we can see below the two regions differ in the expression levels of several genes.

```{r cmp_volcano}
cmp_volcano(cmp, text_y_lim = 20, text_size = 2, text_x_lim = 2.5)
```

We may check the differential expression of these genes using the `cmp_images()` function:

```{r}
markers <- cmp@stat_test %>% 
              dplyr::filter(abs(log2_ratio) > 2.5) %>% 
              dplyr::arrange(log2_ratio) %>%  
              dplyr::filter(-log10(p_values) > 20) %>% rownames()

cmp_images(st_obj_r1_1,
           st_obj_r1_2,
           st_obj_r2_1, 
           st_obj_r2_2,
           feat_list=markers,
           names=c("region_1_1", "region_1_2", "region_2_1", "region_2_2"),
           color_y_strip=c("pink", "pink", "lightblue", "lightblue"),
           color_strip_text_y = "black",
           color_x_strip= c("darkgreen", "darkgreen", "darkred", "darkred", "darkred"),
           colors = viridis::magma(10))
```





