---
title: "Intratumoral dendritic cell"
author: "D. Puthier"
date: "2024-05-14"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Intratumoral dendritic cell–CD4+ T helper cell niches enable CD8+ T cell differentiation following PD-1 blockade in hepatocellular carcinoma

https://zenodo.org/records/7758080

### Loading data

Here we focus on two samples (122_r1, untreated;  1012_r0, cemiplimab).

```{r}
library(stcompr)
set_verb_level(1)
wdir <- "/Users/puthier/Documents/ST_data/dendritic_cells/data"

## Untreated
sto_122_r1_unt <- load_spatial(file.path(wdir,
                                        "122_region_1_detected_transcripts.csv"), 
                              method = "merscope_csv", 
                              sep=",", 
                              bin_size = 25)
sto_122_r1_unt <- rm_controls(sto_122_r1_unt)



sto_1012_r0_cem <- load_spatial(file.path(wdir, "1012_region_0_detected_transcripts.csv"),
                              method = "merscope_csv",
                              sep=",", bin_size = 25)
sto_1012_r0_cem <- rm_controls(sto_1012_r0_cem)



cmp_images(sto_122_r1_unt,
              sto_1012_r0_cem,
            feat_list = "count_sum",
           names=nm)


```

### Distributions

```{r}
# Provide all STGrid object to dist_st()...

nm <- c("sto_122_r1_unt",
        "sto_1012_r0_cem")

dist_st(sto_122_r1_unt,
        sto_1012_r0_cem,
        transform = "log10",
        type = "boxjitter",
        names = nm)


dist_st(sto_122_r1_unt,
        sto_1012_r0_cem,
        transform = "log10",
        names = nm,
        type = "boxjitter",
        normalized=TRUE)

```





### Density

```{r}
## This is extremely (too?) long...
sto_122_r1_unt <- compute_k_ripley(sto_122_r1_unt, rmax = 200, nlarge = 2e6)
sto_1012_r0_cem <- compute_k_ripley(sto_1012_r0_cem, rmax = 200, nlarge = 2e6)
```

### Compare objects

```{r}
st_comp <- stcompr(sto_122_r1_unt, 
                   sto_1012_r0_cem, 
                   name_1 = "sto_122_r0_unt", 
                   name_2 = "sto_1012_r0_cem")
```

### Volcano

```{r}
cmp_volcano(st_comp, text_x_lim = 2, text_y_lim=100, text_size = 4, max.overlaps = 30)
```

### Look at DGE (img + barplot)

```{r}
feat <-  stat_test(st_comp) %>% 
            dplyr::filter(abs(log2_ratio) >3.5) %>% 
            dplyr::filter(-log10(p_values) > 300) %>% 
            dplyr::arrange(log2_ratio, p_values) %>% 
            rownames()

cmp_images(sto_122_r1_unt, 
                   sto_1012_r0_cem, 
                   names= c("sto_122_r0_unt", "sto_1012_r0_cem"), feat_list= feat) 
```

```{r}
cmp_bar_plot(st_comp, features = feat, normalized = TRUE, transform = "None")
```

### Are CD79A... JCHAIN co-localized

```{r}
hc_1012_r0_cem<- hc_tree(sto_1012_r0_cem, class_nb = 24, 
        size = 1, 
        method = "ward.D")
hc_1012_r0_cem
```

```{r}
hc_1012_r0_cem$tree_classes
```
```{r}
# B / plasmablaste ?
cmp_images(sto_122_r1_unt, 
                   sto_1012_r0_cem, 
                   names= c("sto_122_r0_unt", "sto_1012_r0_cem"), feat_list= hc_1012_r0_cem$tree_classes$MOD12) 

# T
cmp_images(sto_122_r1_unt, 
                   sto_1012_r0_cem, 
                   names= c("sto_122_r0_unt", "sto_1012_r0_cem"), feat_list= hc_1012_r0_cem$tree_classes$MOD16)
# B (less mature ?)
cmp_images(sto_122_r1_unt, 
                   sto_1012_r0_cem, 
                   names= c("sto_122_r0_unt", "sto_1012_r0_cem"), feat_list= hc_1012_r0_cem$tree_classes$MOD08)

# Macro 
cmp_images(sto_122_r1_unt, 
                   sto_1012_r0_cem, 
                   names= c("sto_122_r0_unt", "sto_1012_r0_cem"), feat_list= hc_1012_r0_cem$tree_classes$MOD13)
```

### Visualize the gene module/program

```{r}
sto_1012_r0_cem <- compute_module_score(sto_1012_r0_cem, 
                                       modules = hc_1012_r0_cem$tree_classes)
spatial_image(sto_1012_r0_cem, features = meta_names(sto_1012_r0_cem)[-1], ncol = 6)

```

### Identification of cell pops

```{r}
library(clustermole)
markers_list <- 
my_overlaps <- clustermole_overlaps(genes = markers_list, species = "hs")
markers_list <- list()
for(i in 2:length(markers_list)){ 
  try(markers_list[[i]] <- clustermole_overlaps(genes = hc_1012_r0_cem$tree_classes[[i]], 
                             species = "hs"))
}
```




